---
title: 'Summary'
author: "Rebecca Silva"
date: "8/4/2019"
output: 
  pdf_document:
    fig_height: 3
    fig_width: 5
  html_document:
    fig_height: 3
    fig_width: 5
  word_document:
    fig_height: 3
    fig_width: 5
---

```{r, setup, include=FALSE}
knitr::opts_chunk$set(
  tidy=FALSE,     # display code as typed
  size="small")   # slightly smaller font for code
```

## Dipsea Race 

###Data 
```{r}
head(ds)  #created in DipseaData.Rmd
head(ds_NN) #created in NN_Model
```

###Confidence Intervals for Factors
* Basic Idea: Find age-ratio for ages 27, 30, 35. Compare Age-ratio by sex with Age-ratio by sex and section. 
 + bootstrap gam/NN 1000/100 times: caret::train $model = "gam"$ or $model = "nnet"$.
 + sex only model not very strong 
  - For Neural Network best R-Squared ~ .26 (compared with ~ .5 for Sex, Section and Section alone R-Squared ~ .43) 
  - For GAM best R-Squared ~ .27 (compared with ~ .49 for Sex, Section and Section alone R-Squared ~ .42) 
  
####Neural Network Model Bootstrap 

Example of bootstrap fn (Sex and Section model): 
```{r}
boot_nnet<- function (data, testdata, test_scaled, orig_data){
  index = sample( seq_len ( nrow ( data) ), size = nrow( data ) ) 
  train = data[index , ]
  model.nn <- train( Actual_Time ~ Age + Section_NN + Sex_NN,
                  data = train,
                  method = "nnet",
                  trace = FALSE)
  predictions <- predict( model.nn, test_scaled )
  predictions_rescale = (predictions * (max(orig_data$Actual_Time) - 
                                        min(orig_data$Actual_Time)))+ 
    min(orig_data$Actual_Time)
  prediction_data <- cbind(test, predictions_rescale)  

  index <- 1
  time_FR_NN <- c(); time_FI_NN <- c()
  time_MR_NN <- c(); time_MI_NN <- c()
  female <- prediction_data %>% filter(Sex_NN == 0) 
  male <- prediction_data %>% filter(Sex_NN == 1) 
  
  for (age in base_years){ #gets base times (27,30,35)
     time_FR_NN[index] <- unlist(female %>% 
                               filter(Age == age, Section_NN == 0) %>% 
                               select(predictions_rescale))
     time_FI_NN[index] <- unlist(female %>% 
                               filter(Age == age, Section_NN == 1) %>% 
                               select(predictions_rescale))
    time_MR_NN[index] <- unlist(male %>% 
                               filter(Age == age, Section_NN == 0) %>% 
                               select(predictions_rescale))
    time_MI_NN[index] <- unlist(male %>% 
                               filter(Age == age, Section_NN == 1) %>% 
                               select(predictions_rescale))
    
    index <- index + 1
  }
  
FR_time <- unlist(female %>% filter(Section_NN == 0) %>% 
               select(predictions_rescale)) # times for females all ages 
FI_time <- unlist(female %>% filter(Section_NN == 1) %>% 
               select(predictions_rescale))
MR_time <- unlist(male %>% filter(Section_NN == 0) %>% 
               select(predictions_rescale))
MI_time <- unlist(male %>% filter(Section_NN == 1) %>% 
               select(predictions_rescale))

out$factor30_FR = FR_time/time_FR_NN[2]
out$factor30_FI = FI_time/time_FI_NN[2]
out$factor30_MR = MR_time/time_MR_NN[2]
out$factor30_MI = MI_time/time_MI_NN[2]#times divided by time at age 30 

  return( out )
}

#returns list of 4, each has 75 rows of B length vectors
```

After tidying output:
```{r}
head(factors30_SexSectiondf)
head(factors30_Sexdf)
```


#####Visuals
```{r}
require(gridExtra)
SexNN.plot
SexSectionNN.plot
grid.arrange(SexNN.plot, SexSectionNN.plot, ncol = 2)
```

#### GAM Results: 
```{r}
head(factors30_Sexdf)
head(factors30_SexSectiondf)
```

```{r}
SexGAM.plot
SexSectionGAM.plot
grid.arrange(SexGAM.plot, SexSectionGAM.plot, ncol = 2)
```


## US Masters Swimming Data (Top 10)

In Google Doc folder: 'USMS_Top10_SCY.cvs', 'USMS_Top10_LCM.csv', 'USMS_Top10_SCM.csv'



