---
title: "NN_Boot"
author: "Rebecca Silva"
date: "7/11/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This Rmd is my first attempt at bootstraping with the neuralnet() model. This function can take over an hour to build one model and has presented many issues with accuracy. 

Steps in rmd: 
### 1. NN Sex model: Focus on factor age 30 first. 

## 1a. Get CIs for factors 

### 2. NN Sex and Section model: factor 30 

## 2a. compare factors for inv. vs runner

#### 1 ###################################


```{r}
#set.seed(13)

#uses 'scaled' dataset from NN_model.Rmd 


boot_NN_sex <- function(scaled_data, samplesize, testdata, ds_NN, rescaled_testdata){
  index = sample( seq_len ( nrow ( scaled_data ) ), size = samplesize )
  trainNN = scaled_data[index , ]
  NN_sex = neuralnet(Actual_Time ~ Age + Sex_NN , 
                     trainNN, hidden = 2 , linear.output = T, stepmax = 1e7 )
  predict_testNN = predict(NN_sex, testdata) 
  predict_rescale = (predict_testNN * (max(ds_NN$Actual_Time) - min(ds_NN$Actual_Time)))
    + min(ds_NN$Actual_Time) #rescale predicted times
  prediction_data <- cbind(rescaled_testdata, predict_rescale)


  index <- 1
  time_F_NN <- c()
  time_M_NN <- c()
  out <- list()
  female <- prediction_data %>% filter(Sex_NN == 0) 
  male <- prediction_data %>% filter(Sex_NN == 1) 
  
  for (age in base_years){
     time_F_NN[index] <- unlist(female %>% 
                               filter(Age == age) %>% 
                               select(predict_rescale)) 
    time_M_NN[index] <- unlist(male %>% 
                              filter(Age == age) %>% 
                              select(predict_rescale)) 
    index <- index + 1
  }
  
female_time <- unlist(female %>% 
               select(predict_rescale)) #times for females all ages 
male_time <- unlist(male %>% 
               select(predict_rescale))

out$factor30_F = female_time/time_F_NN[2] ; out$factor30_M = male_time/time_M_NN[2] #times divided by time at age 30 

return(out) 
}


factor30_NN <- do(1)*boot_NN_sex(scaled_sex, samplesize, testdata = scaled_newdata, rescaled_testdata = newdata_NN)

### convert to tidy
factor30_NN <- lapply(factor30_NN, function(x){  #make list of dataframes each with 75 rows and B cols 
  as.data.frame(sapply(x,rbind))
  })

### find mean, sd, CI 

summary_factor3_NN <- lapply(factor30_NN, function(x){
  sapply(x, function(x) c( "Stand dev" = sd(x), 
                         "Mean"= mean(x),
                         "LowerCI" = mean(x) - 1.96*sd(x),
                         "UpperCI" = mean(x) + 1.96*sd(x),
                         "Median" = median(x))
  )
})
```


### Visuals 
```{r}
f <- as.vector(factor30_NN$factor30_F) ; m <- as.vector(factor30_NN$factor30_M)
df <- as.data.frame(cbind( Age75, f))
dm <- as.data.frame(cbind( Age75, m))
ggplot() + geom_point( data = df, aes(Age75, f), color = "blue") +
  geom_point( data = dm, aes(Age75, m), color = "red")
#interesting because even relative to their own sex, males deteriote and improve at slower speeds (ie older male needs more headstart relative to 30 male than older female relative to 30 yr old female)
```

### 2. NN Sex and Section 
 Issues I am facing: some model training do not reach that many steps and are poor models for accuracy (about 1 in 10). 

```{r}
samplesize = 0.70 * nrow(ds_NN) 
set.seed(13)

### create test data 
AgeSection <- rep(c(6:80), 4)
SexSection<- c(rep(0, 150), rep(1, 150))
Section <- c(rep(0, 75), rep(1, 75), rep(0, 75), rep(1, 75)) 
testdataSection <- data.frame( AgeSection , SexSection, Section)



### scale test data 
max = apply(testdataSection, 2, max)  #newdata_NN has age and binary sex
min = apply(testdataSection, 2, min)
scaled_testdataSection <- scale(testdataSection, center = min, scale = max- min)

### boot function trains model with neuralnet, then predicts on all combinations, then finds factors. 
boot_NN_Section <- function(scaled_data, samplesize, ds_NN, testdata, rescaled_testdata){
  index = sample( seq_len ( nrow ( scaled_data) ), size = samplesize ) 
  trainNN = scaled_data[index , ]  
  NN_section = neuralnet(Actual_Time ~ Age + Section_NN + Sex_NN, 
                     trainNN, hidden = 1 , linear.output = T, stepmax = 1e7 )
  
  
  predict_testNN = predict(NN_section, testdata)
  # rescale predicted times
  predict_rescale = (predict_testNN * (max(ds_NN$Actual_Time) - 
                                         min(ds_NN$Actual_Time))) + min(ds_NN$Actual_Time)
  # bind rescaled predicted values with all age, sex, and section combinations
  prediction_data <- cbind(rescaled_testdata, predict_rescale)  
  
  # initialize objects for upcoming loop
  index <- 1
  time_FR_NN <- c(); time_FI_NN <- c()
  time_MR_NN <- c(); time_MI_NN <- c()
  # predicted times for females and males 
  female <- prediction_data %>% filter(SexSection == 0)  
  male <- prediction_data %>% filter(SexSection == 1)    
  
  # finds prediction for ages 27, 30, 35 
  for (age in base_years){ 
     time_FR_NN[index] <- unlist(female %>% 
                               filter(AgeSection == age, Section == 0) %>% 
                               select(predict_rescale))
     time_FI_NN[index] <- unlist(female %>% 
                               filter(AgeSection == age, Section == 1) %>% 
                               select(predict_rescale))
    time_MR_NN[index] <- unlist(male %>% 
                               filter(AgeSection == age, Section == 0) %>% 
                               select(predict_rescale))
    time_MI_NN[index] <- unlist(male %>% 
                               filter(AgeSection == age, Section == 1) %>% 
                               select(predict_rescale))
    
    index <- index + 1
  }
  
  # find predictions for each combination 
femaleR_time <- unlist(female %>% filter(Section == 0) %>% 
               select(predict_rescale)) #times for females all ages 
femaleI_time <- unlist(female %>% filter(Section == 1) %>% 
               select(predict_rescale))
maleR_time <- unlist(female %>% filter(Section == 0) %>% 
               select(predict_rescale))
maleI_time <- unlist(female %>% filter(Section == 1) %>% 
               select(predict_rescale))

# finds factors by dividing predicted times by predicted base times (27, 30, 35) 
out$factor30_FR = femaleR_time/time_FR_NN[2] #times divided by time at age 30
out$factor30_FI = femaleI_time/time_FI_NN[2]
out$factor30_MR = maleR_time/time_MR_NN[2]
out$factor30_MI = maleI_time/time_MI_NN[2] 
#divide by time_F_NN[1] and time_F_NN[3] for age 27 and 35 respectively 
return(out) 
}

set.seed(1)
# Run boot function B times with do(B) function (returns list of 4, each list has B columns of vectors length 75)
out <- list()

factor30S_NN <- do(2)*boot_NN_Section(scaled_data = scaled, samplesize, ds_NN, testdata = scaled_testdataSection, rescaled_testdata = testdataSection)

# tidy to dataframe with 4*B (B = # bootstraps) columns and 75 rows with factors 
factor30S_NN <- lapply(factor30S_NN, function(x){  
  as.data.frame(sapply(x,rbind))
  })
```


### Visuals 

```{r}
fr <- as.vector(factor30S_NN$factor30_FR) ; mr <- as.vector(factor30S_NN$factor30_MR)
fi <- as.vector(factor30S_NN$factor30_FI) ; mi <- as.vector(factor30S_NN$factor30_MI)
dfr <- as.data.frame(cbind( Age75, fr)) ; dmr <- as.data.frame(cbind( Age75, mr))
dfi <- as.data.frame(cbind( Age75, fi)) ; dmi <- as.data.frame(cbind( Age75, mi))


ggplot() + geom_point( data = dfr, aes(Age75, fr), color = "blue") +
  geom_point( data = dmr, aes(Age75, mr), color = "red") +
  geom_point( data = dfi, aes(Age75, fi), color = "green") +
  geom_point( data = dmi, aes(Age75, mi), color = "black")

```
