---
title: "Section vs Sex"
author: "Rebecca Silva"
date: "7/22/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(caret) # for train()
library(nnet)  # for nnet()
library(NeuralNetTools)  # for neuralnet()
library(tidyr)
library(ggplot2)
library(mosiac)
library(mgcv)  # for gam()
```
Exploring the train() function which either uses CV or bootstrap to find optimal model. At first, I test with train() without specifying hidden layers for the Nearual network model or basis dimension for the GAM model. 

## Sex and Age, Neural network
```{r}
index = sample( seq_len ( nrow (scaled) ), size = nrow( scaled ) ) 
train = scaled[index , ]
set.seed(23)
model.SexNN <- train( Actual_Time ~ Age + Sex_NN,
                  data = train,
                  method = "nnet",
                  trace = FALSE) 
print(model.SexNN) #best Rsquared = .26, RMSE = .07846
```

##Section, Sex and Age, Neural Network
```{r}
index = sample( seq_len ( nrow (scaled) ), size = nrow( scaled ) ) 
train = scaled[index , ]
set.seed(32)
model.SectionNN <- train( Actual_Time ~ Age + Section_NN,
                  data = train,
                  method = "nnet",
                  trace = FALSE) # add linout = T next time 
print(model.SectionNN)  #best RMSE = .06937, Rsquared = .423
```

## Sex, Section, and age, Neural Network w specifications 
```{r}
## specify boot vs CV and number of reps 
ctrl <- trainControl(method = "boot", number = 10, returnResamp = "all")
# specify number of hidden layers and regularization paraters (decay) to try in train
nnetGrid <-  expand.grid(size = seq(from = 1, to = 3, by = 1),
                        decay = seq(from = 0.0, to = 0.3, by = 0.05))

model.SectionNN <- train( Actual_Time ~ Age + Section_NN + Sex_NN,
                  data = scaled,
                  method = "nnet",
                  metric = "Rsquared",
                  trControl = ctrl,
                  tuneGrid = nnetGrid,
                  trace = FALSE)  
print(model.SectionNN)


model.SectionNN2 <- train( Actual_Time ~ Age + Section_NN + Sex_NN,
                  data = scaled,
                  method = "nnet",
                  trControl = ctrl,
                  tuneGrid = nnetGrid,
                  trace = FALSE) # add linout = T next time 
print(model.SectionNN2)

```

```{r}
# after using train to chose size and decay, we can just train one model using nnet()

nnet_model <- nnet(Actual_Time ~ Age + Section_NN + Sex_NN,
                  data = scaled, size = 2)
print(nnet_model)
predict(nnet_model, test_scaled) # prediction not working
```


## GAM
```{r}
index = sample( seq_len ( nrow ( ds_NN) ), size = nrow( ds_NN ) ) 
train = ds_NN[index , ]
train2 = ds[index,]

set.seed(2)
model.SexGAM <- train( Actual_Time ~ Age + Sex_NN,
                  data = train,
                  method = "gam",
                  trace = FALSE) # add linout = T next time 

print(model.SexGAM) #rsquared= .269 rmse = 0.0776

 
 
set.seed(4)
model.SexGAM2 <-train( Actual_Time ~ Age  + Section_NN,
                  data = train,
                  method = "gam",
                  trace = FALSE)

print(model.SexGAM2) #RMSE = 0.065, Rsquare = .4905
  
model.SexGAM3 <-train( Actual_Time ~ Age + Sex + Section,
                  data = train2,
                  method = "gam",
                  trace = FALSE) 
print(model.SexGAM3) #rsquared = .495

```

Look further into caret::train() using "gam". 
1. See how train() is fitting model without any specification of degrees of freedom or basis dimension. 
2. Look at variation in train() gam models
3. Add specifications?


```{r}
# 1.
modelGAM <- train( Actual_Time ~ Age + Section + Sex,
                  data = ds,
                  method = "gam",
                  trace = FALSE)
print(modelGAM)
summary(modelGAM)
# 2. 
modelGAM1 <- modelGAM
modelGAM2 <- train( Actual_Time ~ Age + Section + Sex,
                  data = ds,
                  method = "gam",
                  trace = FALSE)
summary(modelGAM2) # barely changing 
modelGAM3 <- train( Actual_Time ~ Age + Section + Sex,
                  data = ds,
                  method = "gam",
                  trace = FALSE)

# 3. with specifications

modelGAM <- train( Actual_Time ~ Age + Section + Sex,
                  data = ds,
                  method = "gam",
                  trControl = trainControl(method = "boot", number = 10), 
                  tuneGrid = data.frame(method = "REML", select = FALSE),
                  trace = FALSE)

print(modelGAM)
# train only gives one option of a model to choose from. Not ideal. Therefore want gam() and not train() with 'gam' specification in bootstrap. 
```





