---
title: "Model Fitting"
author: "Rebecca Silva"
date: "June 27, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This rmd is my first attempt at working through gam models. I use the dataset 'ds' and the 'mgvc' package. 

```{r}
library(mosaic)
library("mgcv") # for gam()
```

## "GAM" : for sex and section
```{r}
#want to keep Sex and Section sep. ie do not need Sex_Section var
set.seed(34)
GAM <- mgcv::gam(Actual_Time ~ s(Age, k= 40) + Sex + Section + s(Age, k= 60, by = as.factor(Sex)) + s(Age, k= 60, by = as.factor(Section)) , data = ds)  #.503

plot(GAM, se = T)
summary(GAM) 

mgcv::gam.check(GAM)


#test if sex interaction needed with anova
GAM_minus <- mgcv::gam(Actual_Time ~ s(Age, k= 40) + Sex + Section + s(Age, k= 60, by = as.factor(Section)) , data = ds) 

anova(GAM_minus, GAM, test = "F") #yes, needed
```


###Model for just sex 
```{r}

###### train to find RMSE 
samplesize = 0.70 * nrow(ds_NN)
index = sample( seq_len ( nrow ( ds ) ), size = samplesize )

# Create training and test set
datatrain = ds[ index, ]
datatest = ds[ -index, ]

set.seed(35)
gam_sex <- mgcv::gam(Actual_Time ~ s(Age, k = 40) + s(Age, by = as.factor(Sex), k = 60) + Sex, data = datatrain) 
summary(gam_sex)
plot(gam_sex, se = T)
mgcv::gam.check(gam_sex)

###predict 

pred <- predict(gam_sex, datatest)


y <- ds$Actual_Time[-index ]  #true data for tested

RMSE.GAM <- (sum((y - pred)^2) / nrow(datatest)) ^ 0.5 #16.5
```

# Over fitting? Try smaller k. 
```{r}
set.seed(33) # if want to keep model the same

GAM <- mgcv::gam(Actual_Time ~ s(Age, k= 10) + Sex + Section + s(Age, k= 20, by = as.factor(Sex)) + s(Age, k= 20, by = as.factor(Section)) , method  = 'REML' , data = ds) 

plot(GAM, se = T)
summary(GAM) 

# check degrees of freedom 
mgcv::gam.check(GAM)
```

To make sure our model is not overfitting, we want to look at the variation in prediction between a few of the same model. 

```{r}
GAM <- mgcv::gam(Actual_Time ~ s(Age, k= 10) + Sex + Section + s(Age, k= 20, by = as.factor(Sex)) + s(Age, k= 20, by = as.factor(Section)) , method  = 'REML' , data = ds) 

# create test data 
Age <- rep(c(6:80), 4)
SexGAM <- c(rep("M", 150), rep("F", 150))
SectionGAM<- c(rep("Invitational", 75), 
             rep("Runner", 75), rep("Invitational", 75), rep("Runner", 75)) 
testGAM2 <- data.frame( Age , "Sex" = SexGAM, "Section" = SectionGAM)
testdata = testGAM2

# predict on test
GAM1 <- GAM
set.seed(1)
GAM1 <- mgcv::gam(Actual_Time ~ s(Age, k= 10) + Sex + Section + s(Age, k= 20, by = as.factor(Sex)) + s(Age, k= 20, by = as.factor(Section)) , method  = 'REML' , data = ds) 
pred1 <- predict( GAM1, testdata )
set.seed(2)
GAM2 <- mgcv::gam(Actual_Time ~ s(Age, k= 10) + Sex + Section + s(Age, k= 20, by = as.factor(Sex)) + s(Age, k= 20, by = as.factor(Section)) , method  = 'REML' , data = ds) 
pred2 <- predict( GAM2, testdata )

GAM3 <- mgcv::gam(Actual_Time ~ s(Age, k= 10) + Sex + Section + s(Age, k= 10, by = as.factor(Sex)) + s(Age, k= 10, by = as.factor(Section)) , method  = 'REML' , data = ds) 
pred3 <- predict( GAM3, testdata )

# look at variation
pred_data <- cbind(testdata, pred1, pred2, pred3) 

# no variation. 
summary(GAM1); summary(GAM2); summary(GAM3)
```

# Even smaller k? 
Possible problem with even smaller k: gam.check() says that 'k-index' <1 may indicate k is too small. However it says, "especially when edf is close to k" which is not the typical case here (below). 
```{r}
GAM <- mgcv::gam(Actual_Time ~ s(Age, k= 10) + Sex + Section + s(Age, k= 10, by = as.factor(Sex)) + s(Age, k= 10, by = as.factor(Section)) , method  = 'REML' , data = ds) 

gam.check(GAM)

# create 3 new models and predict on test
GAM1 <- GAM
pred1 <- predict(GAM1, testdata)

set.seed(10)
GAM2 <- mgcv::gam(Actual_Time ~ s(Age, k= 10) + Sex + Section + s(Age, k= 10, by = as.factor(Sex)) + s(Age, k= 10, by = as.factor(Section)) , method  = 'REML' , data = ds) 
pred2 <- predict( GAM2, testdata )
set.seed(3)
GAM3 <- mgcv::gam(Actual_Time ~ s(Age, k= 10) + Sex + Section + s(Age, k= 10, by = as.factor(Sex)) + s(Age, k= 10, by = as.factor(Section)) , method  = 'REML' , data = ds) 
summary(GAM3)
pred3 <- predict( GAM3, testdata )

# look at variation
pred_data2 <- cbind(testdata, pred1, pred2, pred3) # better variation 
```

Try sampling data to avoid overfitting. 
```{r}
samplesize = 0.85 * nrow(ds)
index = sample( seq_len ( nrow ( ds ) ), size = samplesize )
train = ds[index , ]


# create 3 new models and predict on test
g1 <- mgcv::gam(Actual_Time ~ s(Age, k= 10) + Sex + Section + s(Age, k= 10, by = as.factor(Sex)) + s(Age, k= 10, by = as.factor(Section)) , method  = 'REML' , data = train) 
pred1 <- predict(g1, testdata)

index = sample( seq_len ( nrow ( ds ) ), size = samplesize )
train = ds[index , ]
g2 <- mgcv::gam(Actual_Time ~ s(Age, k= 10) + Sex + Section + s(Age, k= 10, by = as.factor(Sex)) + s(Age, k= 10, by = as.factor(Section)) , method  = 'REML' , data = train) 
pred2 <- predict( g2, testdata )


# look at variation
pred_data2 <- cbind(testdata, pred1, pred2) # better variation 
```



